---
title: "VGSales Analysis aand Charts"
output: html_document
date: "2025-03-31"
runtime: shiny
---
```{r include = FALSE}

##install.packages("openxlsx")
##install.packages("rmarkdown")
##install.packages("shiny")

library(openxlsx)
library(ISLR)
library(ggplot2)
library(tidyverse)
library(rmarkdown)
library(shiny)

## Remove any unfinished data (Ones with NA)
vgsales <- read.csv('vgsales.csv')
datFrame <- data.frame(vgsales)
datFrame[datFrame=="N/A"] <- NA
datFrame <- na.omit(datFrame)

## Set data to desired types
str(datFrame)
datFrame$Genre <- as.factor(datFrame$Genre)
datFrame$Publisher <- as.factor(datFrame$Publisher)
datFrame$Platform <- as.factor(datFrame$Platform)
datFrame$Year <- as.numeric(datFrame$Year)
str(datFrame)

## Looking at interests
par(mfrow=c(1,1))
summary(datFrame$Publisher)
```

```{r echo=FALSE}
plot(datFrame$Genre, col = "#FF6666")

summary(datFrame$Year)
summary(datFrame$Platform)
summary(datFrame$Genre)
```

## Plot of Genre per Platform count

```{r Genre Per Platform, echo=FALSE}
ggplot(datFrame) + geom_bar(aes(x=datFrame$Platform, fill=factor(datFrame$Genre)), position = "dodge") + theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
```


## Plot of Overall Consumption of products
```{r Consumer Data , echo=FALSE}
ggplot(datFrame) + geom_bar(aes(x=datFrame$Year, fill=factor(datFrame$Genre)), position = "dodge") + theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
ggplot(datFrame) + geom_bar(aes(x=datFrame$Year)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
```

## Plot of Platform per Year Count
```{r Platform Per Year, echo=FALSE}

ggplot(datFrame) + geom_bar(aes(x=datFrame$Year, fill=factor(datFrame$Platform)), position = "dodge") + theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
```

## Overall Platform count
```{r Platform Count, echo=FALSE}


ggplot(datFrame) + geom_bar(aes(x=datFrame$Platform)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels
```

## Timeline of all platforms
```{r Timeline, echo=FALSE}


ggplot(datFrame, aes(x = datFrame$Platform, y = datFrame$Year)) + geom_tile() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(datFrame, aes(x = Platform, y = Year)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

## Games Produced For Platforms Over Time
```{r View Platform Data, echo=FALSE}

ui <- fluidPage(

  sidebarLayout(
    sidebarPanel(
      selectInput("platform", "Choose a platform:", 
                  choices = unique(datFrame$Platform), 
                  selected = unique(datFrame$Platform)[1])
    ),
    
    mainPanel(
      plotOutput("platformPlot")
    )
  )
)

server <- function(input, output) {
  # Reactive dataset
  PlatformTable <- reactive({
    datFrame %>% filter(Platform == input$platform)
  })
  
  # Render plot dynamically
  output$platformPlot <- renderPlot({
  PlatformTable() %>%
    group_by(Year) %>%
    summarise(Game_Count = n()) %>%  # Number of games per year
    ggplot(aes(x = Year, y = Game_Count)) + 
    geom_bar(stat = "identity", fill = "steelblue") +  # Show actual counts
    labs(title = paste("Number of Games for", input$platform), x = "Year", y = "Number of Games") +
    scale_x_continuous(breaks = seq(min(PlatformTable()$Year), max(PlatformTable()$Year), by = 1)) +  # Whole numbers
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +  # Make y-axis readable
    theme_minimal()
})
}

shinyApp(ui, server)

```

## Sales for For Games on Selected Platform (Year Made)
```{r Platform Sales, echo=FALSE}
ui <- fluidPage(

  sidebarLayout(
    sidebarPanel(
      selectInput("salesPlatform", "Choose a platform:", 
                  choices = unique(datFrame$Platform), 
                  selected = unique(datFrame$Platform)[1])
    ),
    
    mainPanel(
      plotOutput("salesPlot"),
      DT::dataTableOutput("salesTable")  # Table for sales
    )
  )
)

server <- function(input, output) {
  # Reactive dataset
  SalesTable <- reactive({
    datFrame %>%
      filter(Platform == input$salesPlatform) %>%
      group_by(Year) %>%
      summarise(Total_Sales = sum(Global_Sales, na.rm = TRUE))  # Sum sales per year
  })
  
  # Render sales plot dynamically
  output$salesPlot <- renderPlot({
    SalesTable() %>%
      ggplot(aes(x = Year, y = Total_Sales)) + 
      geom_bar(stat = "identity", fill = "darkorange") + 
      labs(title = paste("Total Sales for", input$salesPlatform), x = "Year", y = "Total Sales (millions)") +
      scale_x_continuous(breaks = seq(min(SalesTable()$Year), max(SalesTable()$Year), by = 1)) +  # Lock x-axis
      scale_y_continuous(labels = scales::comma) +  # Format 
      theme_minimal()
  })
  
  # Render sales data table
  output$salesTable <- DT::renderDataTable({
    SalesTable() %>%
      mutate(Total_Sales = round(Total_Sales, 2))  # Round
  })
}

shinyApp(ui, server)
```

## Sales of Genres Over Time (Based on Year Released)
```{r Sales of Genres Over Time, echo=FALSE}

library(shiny)
library(ggplot2)
library(dplyr)

# Ensure Year is number
datFrame$Year <- as.numeric(datFrame$Year)

global_min <- 0
global_max <- max(datFrame$Global_Sales, na.rm = TRUE) * 2  # Lock scaling

ui <- fluidPage(

  fluidRow(
    column(3,  
      checkboxGroupInput("selectedGenres", 
                         "Choose Genres to Display:", 
                         choices = unique(datFrame$Genre), 
                         selected = unique(datFrame$Genre))  # By default, all genres are selected
    ),
    
    column(9,  
      plotOutput("genreSalesPlot")  
    )
  )
)

server <- function(input, output) {
  output$genreSalesPlot <- renderPlot({
    # Filter dataset for selected genres
    GenreSales <- datFrame %>%
      filter(Genre %in% input$selectedGenres) %>%
      group_by(Year, Genre) %>%
      summarise(Total_Sales = sum(Global_Sales, na.rm = TRUE), .groups = 'drop')

    # Ensure y-axis doesn't cut off points
    max_genre_sales <- max(GenreSales$Total_Sales, na.rm = TRUE) * 1.2  # more space for points

    ggplot(GenreSales, aes(x = Year, y = Total_Sales, color = Genre)) +
      geom_line(size = 1.2) + 
      geom_point(size = 3) +  
      labs(title = "Sales Trend for Selected Genres Over Time",
           x = "Year",
           y = "Total Sales (millions)") +
      scale_y_continuous(limits = c(global_min, max(global_max, max_genre_sales)), labels = scales::comma) +  
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1)  
      )
  })
}

shinyApp(ui, server)

```

## Total Sales for Each Genre Across Platforms
```{r Sales by Genre and Platform, echo=FALSE}


ui <- fillPage(
  
  fluidRow(
    column(3, 
      selectInput("selectedPlatform", "Choose a Platform:", 
                  choices = unique(datFrame$Platform), 
                  selected = unique(datFrame$Platform)[1])
    )
  ),
  
  fluidRow(
    column(12, plotOutput("genreSalesPlot", height = "400px")) ## Fixed height
  )
  

)

server <- function(input, output) {
  output$genreSalesPlot <- renderPlot({
    # Filter dataset by the selected platform
    GenreSales <- datFrame %>%
      filter(Platform == input$selectedPlatform) %>%
      group_by(Genre) %>%
      summarise(Total_Sales = sum(Global_Sales, na.rm = TRUE)) %>%
      ungroup()
    
    # Create bar plot
    ggplot(GenreSales, aes(x = Genre, y = Total_Sales, fill = Genre)) +
      geom_bar(stat = "identity") +
      labs(title = paste("Total Sales by Genre for", input$selectedPlatform),
           x = "Genre",
           y = "Total Sales (millions)") +
      scale_y_continuous(labels = scales::comma) +  # Format 
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=1), # Rotate labels
      plot.margin = margin(10, 10, 150, 10)) +  # Increase bottom margin
      coord_cartesian(clip = "off") +
      expand_limits(y = max(GenreSales$Total_Sales) * 1.1) # Ensure labels are not cut off
  }
)
}

shinyApp(ui, server)

```



## Sales Distribution by Genre

```{r Sales Distribution by Genre}

ui <- fillPage(

    fluidRow(
    column(3, 
        selectInput("selected_platform", "Choose a Platform:", 
          choices = unique(datFrame$Platform), 
            selected = "Wii")
    )
  ),
  
  fluidRow(
    column(12, plotOutput("pieChart")) ## Fixed height
  )

)

server <- function(input, output) {
    # Reactive dataset
    filtered_data <- reactive({
        datFrame %>%
            filter(Platform == input$selected_platform) %>%
            group_by(Genre) %>%
            summarise(total_sales = sum(Global_Sales, na.rm = TRUE))
    })
    
    # Render Pie Chart
    output$pieChart <- renderPlot({
        data <- filtered_data()
        pie(data$total_sales, labels = data$Genre, 
            main = paste("Sales Distribution for", input$selected_platform))
    })
}

shinyApp(ui, server)

```


## Linear regressions

### Genre Linear Regression Model

```{r Genre Linear Regression Model}


## QQ plots are right sqewed
lm_Genre <- lm(datFrame$Global_Sales ~ datFrame$Genre, data = datFrame)
summary(lm_Genre)

```
Any entry with a p-value of < 0.05 is significant

```{r include = FALSE}
plot(datFrame$Genre, datFrame$Global_Sales)
abline(lm_Genre, col = "blue")
plot(lm_Genre)
```


### Platform Linear Regression Model

```{r Platform Linear Regression Model}

lm_Platform <- lm(datFrame$Global_Sales ~ datFrame$Platform, data = datFrame)
summary(lm_Platform)

```
Any entry with a p-value of < 0.05 is significant


```{r include = FALSE}

plot(datFrame$Platform, datFrame$Global_Sales)
abline(lm_Platform, col = "blue")
plot(lm_Platform)


lm_Year <- lm(datFrame$Global_Sales ~ datFrame$Year, data = datFrame)
summary(lm_Year)
plot(datFrame$Year, datFrame$Global_Sales)
abline(lm_Year, col = "blue")
plot(lm_Year)
```
